Phylogenetics and the human microbiome.

### 2013-09-25 sbreview.tex
Roughed out the data section.

@@

As described above ``metagenome" means that data is sheared randomly.
This type of data has a complication above that taken from 16S survey data, which is that the genetic region sequenced is unknown in addition to the organism it was taken from.

Some use metagenomic data as an expanded marker gene library.
\citep{kembel2011phylogenetic}
AMPHORA, Phylosift.
One approach to using metagenomic data is simply to subset to genes that are found in a large proportion of micro-organisms.
This is the approach taken by \citep{von2007quantitative,stark2010mltreemap} and Jonathan Eisen \citep{wu2008amphora} (in preparation).
Because of the diversity of gene repertoire in microbes, the gene sets may have only limited overlap with one another, and even the largest collection of genes in these databases only recruits around 1 percent of a metagenome.

An alternative approach is to use metagenomic reads directly for analysis.
The authors of the metAMOS pipeline \citep{treangen2013metamos} report speedups and much higher accuracy when reads are assembled before they are classified.

If not used as an expanded marker gene panel then typically used to infer information about metabolic capacity (cite Humann) rather than make statements of an evolutionary nature.

\subsection{Whole genomes}
Whole-genome sequencing used for microbial outbreak tracking \citep{snitkin2012tracking}.
Now from culture, but next we will see single-cell sequencing.

\subsection{Data availability and future}
reference genomes
HMP DACC.

HMP nature paper
\citep{methe2012framework}

>>>>

Lots of survey, metagenome, and whole-genome sequencing data has been generated by the Human Microbiome Project \citep{methe2012framework} and is available on its website\footnote{\url{http://www.hmpdacc.org/}}.

@@

As described above ``metagenome" means that data is sheared randomly.
This type of data has a complication above that taken from 16S survey data, which is that the genetic region sequenced is unknown in addition to the organism it was taken from.
However, because it does proceed through an amplification step, it does not have the same primer biases as a marker gene survey, although sequencing is known to have biases.

It is possible to use metagenomic data as an expanded set of marker genes.
That is, one can use 16S reads that appear in the metagenome as well as other ``core" genes present in a large proportion of micro-organisms that are expected to follow the same evolutionary path \citep{von2007quantitative,wu2008amphora,stark2010mltreemap,kembel2011phylogenetic}.
Because of the diversity of gene repertoire in microbes, the gene sets may have only limited overlap with one another, and even the largest collection of genes in these databases only recruits around 1 percent of a metagenome.

Metagenomic data can be used to infer information about metabolic capacity \cite{greenblum2012metagenomic,abubucker2012metabolic}.
The authors of the metAMOS pipeline \citep{treangen2013metamos} report speedups and much higher accuracy when reads are assembled before they are classified.
The assembly of complete genomes from metagenomes, once limited to samples with a very small number of organisms \citep{baker2010enigmatic}, is now becoming feasible for more diverse populations with improved sequencing technology and computational approaches \cite{howe2012assembling,pell2012scaling,iverson2012untangling,emerson2012metagenomic,podell2013assembly}.

\subsection{Whole genomes}
Whole-genome sequencing from culture is currently being used for microbial outbreak tracking \citep{koser2012rapid,snitkin2012tracking}.
The Food and Drug Administration maintains GenomeTrakr, an openly accessible database of whole genome sequences from culture\footnote{\url{http://www.fda.gov/Food/FoodScienceResearch/WholeGenomeSequencingProgramWGS/}}.
This data will become common for unculturable organisms as single-cell sequencing \citep[reviewed in][]{kalisky2011single} becomes common.
